---
title: "DATA 607 - Project 4"
author: "Jered Ataky, Magnus Skonberg"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    df_print: paged
    smooth_scroll: yes
    toc_collapsed: yes
    toc_depth: 5
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
    code_folding: "hide"
---

```{r echo = FALSE}
knitr::opts_chunk$set(eval = TRUE, results = TRUE, fig.show = "asis", message = FALSE)
```


```{r load-packages, include=FALSE}
library(tidyverse)
library(knitr)
library(R.utils)
library(DT)
library(tm)
library(wordcloud)
library(data.table)
```

### Background

The focus of this project is **document classification**. 

For this project, we will start with a corpus dataset, unzip our data, generate a training model that we'll then use to predict the class of new documents (those withheld from the training set or taken from another source), and then analyze the accuracy of our predictive classifier.

### Download Data

We lean on the R.utils library to *automatically* download, bunzip, extract the contents of tar archive into our "emails" directory, and then create a corresponding list of file names from the spam and ham emails available on [spamassassin](https://spamassassin.apache.org/old/publiccorpus/):
```{r}
#Download, bunzip, and untar spam_2 files into "emails" directory
#download.file("http://spamassassin.apache.org/old/publiccorpus/20050311_spam_2.tar.bz2", destfile= "20050311_spam_2.tar.bz2")
#bunzip2("20050311_spam_2.tar.bz2")
#untar("20050311_spam_2.tar", exdir="emails")

#Create corresponding list of file names for spam_2 and exclude cmds file
if (file.exists("emails\\spam_2\\cmds")) file.remove("emails\\spam_2\\cmds")
spam_list = list.files("emails\\spam_2\\") 

#Download, bunzip, and untar easy_ham files into "emails" directory
#download.file("http://spamassassin.apache.org/old/publiccorpus/20030228_easy_ham.tar.bz2", destfile="20030228_easy_ham.tar.bz2")
#bunzip2("20030228_easy_ham.tar.bz2")
#untar("20030228_easy_ham.tar", exdir = "emails")

#Create corresponding list of file names for easy_ham and exclude cmds file
if (file.exists("emails\\easy_ham\\cmds")) file.remove("emails\\easy_ham\\cmds")
ham_list = list.files("emails\\easy_ham\\")

```


#### Get the insights


Check the length and the contents

```{r}

length(spam_list)

```

```{r}
length(ham_list)
```


Data frame

```{r}
# Build data frame

df_mails <- tibble()
df_mails_folders <- c("emails\\spam_2\\", "emails\\easy_ham")
df_mails_types <- c("spam", "ham")

```


```{r}


# For loop to feed the data frame

for (i in 1: length(df_mails_folders))
  
  {
        tag <- df_mails_types[i]
        
        l <- tibble(file = dir(df_mails_folders[i],  full.names = TRUE)) %>% 
          
          mutate(messages = map(file, read_lines)) %>%
          
                transmute(id = basename(file), tag = tag, messages) %>%
                unnest(messages)
                df_mails<- bind_rows(df_mails, l)
 }
```



```{r}

# Build the table with only tag and messages
# Concatenate the message to be as a single string for a given id

new_df <- df_mails[!duplicated(df_mails$id), ]
new_df[, 'messages'] <- aggregate (messages~id, data = df_mails, toString) [,2]

head(new_df)

```


```{r}

# Subset the data frame with only tag and messages 

df_final <- new_df %>%
  select(tag, messages)

dim(df_final)
```

We now character vectors of file names available as spam_list and ham_list. From these lists, we want to create a corpus of training and then test documents. In order to do so ...

### Creation of Corpus


#### Creating corpus

Writing

```{r}

# Create corpus from all the messages

text_corpus <- Corpus(VectorSource(df_final$messages))
print(text_corpus)

```


#### Cleaning corpus

Writing

```{r}

# remove punctuation
clean_corpus <- tm_map(text_corpus, removePunctuation)

clean_corpus<- tm_map(text_corpus,content_transformer(gsub), pattern="\\W",replace=" ")
removeURL <- function(x) gsub("http^\\s\\s*", "", x)%>% 
clean_corpus <- tm_map(clean_corpus, content_transformer(removeURL))
# remove numbers
clean_corpus <- tm_map(clean_corpus, removeNumbers)


## remove stop words
clean_corpus <- tm_map(clean_corpus, removeWords, stopwords())
## remove whitespace
clean_corpus <- tm_map(clean_corpus, stripWhitespace)

#translate all letters to lower case
clean_corpus <- tm_map(clean_corpus, tolower)

```

#### Creation of Term Document Matrix

```{r}
# Corpus insights

text_final <- TermDocumentMatrix(clean_corpus)
text_final

```

```{r}
# Reduce sparsity 

text_final <- text_final %>% 
  removeSparseTerms(1-100/length(corpus))

inspect(text_final)

```


### Preprocessing of Corpus

Writing

```{r}
#code
```

#### Prepare Labels and Features for Matrix

Writing

```{r}
#code
```

#### Create Training and Test Data

Writing

```{r}
#code
```

### Run the Model

Writing

```{r}
#code
```

### Test the Model

Writing

```{r}
#code
```

### Analyze

Write regarding the approach we took, any interesting insights gained over the course of the project or specific to the ham/spam emails, and provide feedback regarding how we might have improved / had a smoother approach.

