---
title: "DATA 607 - Project 3"
author: "Team Silver Fox"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    code_folding: hide
    df_print: paged
    highlight: tango
    #number_sections: yes
    smooth_scroll: yes
    theme: united
    toc_collapsed: yes
    toc_depth: 5
    toc_float: yes
  pdf_document:
    extra_dependencies:
      -geometry
      -multicol
      -multirow
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  word_document:
    toc: year
    toc_depth: "5"
theme: lumen
number_sections: yes
toc_depth: 3
---


```{r load-packages, include=FALSE}
library(tidyverse)
library(openintro)
library(dplyr)
library(tidyr)
library(RCurl)
library(ggplot2)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### BACKGROUND

> What are the most valued data science skills?

Answering this question formed the crux of our project. 

We collaborated in understanding the question at hand, brainstorming an approach and where we might pull data from, deciding on a dataset and then acquiring, tidying & transforming, visualizing & analyzing this dataset before ultimately coming to the conclusion we'll present later.

*Team Silver Fox*  collaborated virtually on Slack, Google Docs, Google Meet, and Github. There were numerous adaptations to our approach along the way and we ultimately decided on a "twice sifted" approach to answering the question above.

Without further adieu, let’s introduce *Team Silver Fox*:

* **Dan Rosenfeld** – [Github link]
* **Magnus Skonberg** – [Github](https://github.com/Magnus-PS/)
* **Mustafa Telab** – [Github](https://github.com/mtelab1)
* **Josef Waples** - [Github](https://github.com/hachiko1024)

--------------------------------------------------------------------------------

#### DATA SOURCE(S)

We identified our sources of data as the following (cited APA-style below):

1. Kaggle. (2018). **2018 Kaggle ML & DS Survey** [Data file]. Retrieved from https://www.kaggle.com/kaggle/kaggle-survey-2018?select=multipleChoiceResponses.csv

2. Jeff Hale. (2018). **The Most in Demand Skills for Data Scientists** [Data file]. Retrieved from https://www.kaggle.com/discdiver/the-most-in-demand-skills-for-data-scientists/log?select=ds_job_listing_software.csv

For more background on Data Source #1:

* The **2018 Kaggle ML & DS Survey** dataset had (3) .csv’s: the Survey Schema, Free Form Responses, and Multiple Choice Responses. 

* *The Survey Schema* contained questions and explanations of response exemptions. While we do not plan on analyzing this .csv, it was useful pre-screening questions for pertinence. 

* *The Free Form Responses* contained text answers submitted by survey takers when their response was not covered by multiple choice options. If we deem this set valuable / unique for purposes of analysis, we will draw on it. Otherwise, it may be disregarded as well.

* *The Multiple Choice Responses* contain categorical variables submitted by respondents. This is where our “gold” lies and likely where the brunt of our data tidying, transforming, and analysis efforts will go.

--------------------------------------------------------------------------------

#### APPROACH

We started our search by utilizing Kaggle’s self proclaimed “most comprehensive dataset available on the state of ML and data science”, pulling what we deemed useful from this set.

From our findings of the general dataset, we then applied a “value filter” (ie. only considering responses above a certain income, education, or experience level) to see how the general perspective of essential skills differed from that of a more seasoned (or market-valued) perspective.

At this point, we then transitioned to comparing data science skills deemed as valuable in Kaggle's 2018 survey with the most in-demand skills for data scientists across multiple job sites. This "pro level" dataset (also from 2018) was prepared by Jeff Hale, and drawn on to paint the picture of contrast (if there was any) between the skills data scientists deem as most valuable (ie. those doing the work) vs. those that employers hire for (ie. hiring managers).

By contrasting our sets via different techniques and visualizations, our findings were twice sifted. Hence the mention of a "twice sifted" approach in the *Background* section.

Our general approach is outlined below:

1. **Pre-screen:** after viewing the initial Kaggle **2018 Kaggle ML & DS Survey** multiple choice response dataset and seeing that we would not be able to access this dataset via Github due to its size (39+ MB), we deemed “pre-screening” essential. We reviewed all 50 questions from the SurveySchema.csv file and kept only those applicable to the question at hand. Once this dataset had been “pre-screened” it was uploaded to Github along with Jeff Hale’s general and software skill csv files (for later comparison).
2. **Acquire data via reading .csv's:** get the URL, read each .csv from Github (in its raw form), and put the data into tabular form before moving on to tidying and transforming.
3. **Tidy and transform:** by leaning on the dplyr() library and numerous useful R functions, we extracted a table of questions and a table of answers prior to extracting relevant keywords and counts (for later visualization). Our T&T process was heavy on the **2018 Kaggle ML & DS Survey** data and light for the **The Most in Demand Skills for Data Scientists** data. 
4. **Visualize and analyze**: once our data was in a more "digestible" form we set out to generate insightful plots regarding the most valuable **software skills** and **general skills**. We interpreted the general v. value-filtered survey data, and then introduced plots for the **The Most in Demand Skills for Data Scientists** data. The purpose of this tiered approach was to sift for additional insight (if there was any to be found).
5. **Conclude**: after applying our tiered approach, we drew insight from the story our data told and came to a conclusion. More on this later :)

--------------------------------------------------------------------------------

### ACQUIRE DATA

First, we read our .csv's from Github (in their raw form) and then converted these datasets into tabular form. Our approach was to create and utilize a function to get corresponding URLs and read the csv files from these URLs. This was not the most line-efficient strategy but it was a unique way of extracting our csv's which some may deem a "rinse and repeat" process:

```{r}
#Store URLs
url1 <- "https://raw.githubusercontent.com/Magnus-PS/CUNY-SPS-DATA-607/Project-3/pertinent_mc_resps.csv"
url2 <- "https://raw.githubusercontent.com/Magnus-PS/CUNY-SPS-DATA-607/Project-3/ds_software.csv"
url3 <- "https://raw.githubusercontent.com/Magnus-PS/CUNY-SPS-DATA-607/Project-3/ds_general.csv"
#Function to get URLs and read in corresponding .csv files
url_to_dataframe_function <- function(x, y, z) {
  url <- getURL(url1)
  data1 <- read.csv(text = url)
  
  url2 <- getURL(url2)
  data2 <- read.csv(text = url2)
  
  url3 <- getURL(url3)
  data3 <- read.csv(text = url3)
  
  return(list(data1, data2, data3))
  }
#Call to function (defined above)
dataframes <- url_to_dataframe_function(url1, url2, url3)
#Assignment of values
for (i in seq(dataframes))
  assign(paste0("data", i), dataframes[[i]])
#Conversion to dataframe and display for 2018 Kaggle ML & DS Survey multiple choice data and then The Most in Demand Skills for Data Scientists software and general skills
mc_resps <- as_tibble(data1)
tech_skills <- as_tibble(data2)
genl_skills <- as_tibble(data3)
```

--------------------------------------------------------------------------------

### TIDY & TRANSFORM

Once we had our data at hand, we set out to make sense of this data and put it in a more digestible (normalized) form for later visualization and analysis.

For this phase of the project, we added to the more ordinary of R functions with our use of the dplyr() library. We removed columns and rows, renamed columns, transposed our data into long form, gathered essential columns into different data frames, extracted relevant counts, and created "light" visualizations meant to provide insight as to what v. wasn't essential for the next phase. 

Our T&T process was heavy on the **2018 Kaggle ML & DS Survey** data and light for the **The Most in Demand Skills for Data Scientists** data, and took the following form: 


#### SURVEY QUESTIONS & ANSWERS

We created a table of questions by extracting unique questions from the existing dataset, compressing repeated questions from multiple columns into one, renaming column titles for the same format (ie. Q4), transposing for a long dataframe, and replacing question numbers with question IDs.

Thus transforming a 2 x 43 subsection of the dataset into a 10 x 2 table:
```{r}
#Extract all unique questions into questions table
questions <- mc_resps[1,]
#Compress questions with multiple columns into one
comp_qs <- questions[ -c(6:12,14:31,33,35,38:43) ]
#Rename column titles for same format (ie. Q4)
names(comp_qs)[names(comp_qs) == "Q11_Part_1"] <- "Q11"
names(comp_qs)[names(comp_qs) == "Q16_Part_1"] <- "Q16"
names(comp_qs)[names(comp_qs) == "Q34_Part_1"] <- "Q34"
#Transpose for a long dataframe
qs_long <- gather(comp_qs, "Q #", "Question", 1:10, factor_key=TRUE)
#Replace the question number with a question ID (ie. Q4 --> 1)
qs_long$Q_ID <- seq.int(nrow(qs_long)) 
qs_long <- qs_long[ -c(1) ]
qs_final <- qs_long[c("Q_ID", "Question")]
qs_final
```
The table above notes the 10 questions we considered in exploring the **2018 Kaggle ML & DS Survey**. While we did not directly tie these questions in with our visualizations (later on), we did refer to these questions for keywords to ensure a consistent interpretation of the data at hand.

We then created a table of answers and further widdled our dataset down by removing all columns with "OTHER" labels or non-valuable entries and removing 1st row question statements to create an answers table that would be drawn from for tables, plots, and analysis later on.

A small yet foundational step.

```{r}
#Further widdle our dataset down: remove all columns with "OTHER" labels and non-valuable entries, remove 1st row (question statement)
answers <- mc_resps[ -c(11:12,29:31,33,35,43) ]
answers <- answers[-1,]
```

Once we'd laid the foundation, we set out to draw counts from this table of answers for software skills, general skills, and the value filter ...

#### SURVEY SOFTWARE SKILLS

Next, we explored the **2018 Kaggle ML & DS Survey** dataset further and set out to create tables of unique variables related to SOFTWARE SKILLS: 

```{r}
#Software language "regular basis" (Q16) count
##Gather associated answers into one column, cut out irrelevant columns, then count
reg <- gather(answers, key = "Q16_A_ID", value = "Q16_Answer", Q16_Part_1, Q16_Part_2, Q16_Part_3, Q16_Part_4, Q16_Part_5, Q16_Part_6, Q16_Part_7, Q16_Part_8, Q16_Part_9, Q16_Part_10, Q16_Part_11, Q16_Part_12, Q16_Part_13, Q16_Part_14, Q16_Part_15, Q16_Part_16)
reg <- reg[ -c(1:19) ]
reg <- reg %>% count(Q16_Answer, sort = TRUE)%>% rename(reg = "n")
reg <- reg [-1,] #remove empty row

#Software language "use most often" (Q17) count
mo <- answers %>% count(Q17, sort = TRUE)%>% rename(mo = "n")
mo <- mo[-1,] #remove empty row

#Software language "recommend to aspiring DSs" (Q18) count
rec <- answers %>% count(Q18, sort = TRUE)%>% rename(rec = "n")
rec <- rec[-2,] #remove empty row
```

#### SURVEY GENERAL SKILLS

We then applie the same methodology to create a table for unique variables for GENERAL SKILLS from the **2018 Kaggle ML & DS Survey** data: 

```{r}
#Machine learning (Q10) count - very specific. Programming languages are a gateway to this direction.
#ml_count <- answers %>% count(Q10, sort = TRUE)
#ml_count

#Daily activities (Q11) count - NOT USEFUL / UNIQUE
##Gather associated answers into one column, cut out irrelevant columns, then count
# all_daily <- gather(answers, key = "Q11_A_ID", value = "Q11_Answer", Q11_Part_1, Q11_Part_2, Q11_Part_3, Q11_Part_4, Q11_Part_5, Q11_Part_6)
# all_daily <- all_daily[ -c(1:29) ]
# all_daily %>% count(Q11_Answer, sort = TRUE)

#Software activity (Q23) count
sw <- answers %>% count(Q23, sort = TRUE)
sw <- sw[-2,] #remove empty row
sw$Q23 <- as.character(sw$Q23)
sw$Q23[sw$Q23 == "50% to 74% of my time"] <- "50%-74%"
sw$Q23[sw$Q23 == "25% to 49% of my time"] <- "25%-49%"
sw$Q23[sw$Q23 == "1% to 25% of my time"] <- "1%-25%"
sw$Q23[sw$Q23 == "75% to 99% of my time"] <- "75%-99%"
sw$Q23[sw$Q23 == "0% of my time"] <- "0%"
sw$Q23[sw$Q23 == "100% of my time"] <- "100%"

#Visualize software activity
ggplot(sw, aes(x = Q23, y = n)) +
  geom_col(fill = "#00abff") +
  labs(title = "Data Science Survey", subtitle = "% of school / workday actively coding", x = "", y = "n") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"), plot.subtitle = element_text(hjust = 0.5)) +
  labs(caption = "2018") +
  scale_x_discrete(limits= c("0%", "1%-25%", "25%-49%","50%-74%", "75%-99%", "100%"))

#% use time (Q34) count - NOT USEFUL / UNIQUE
##Gather associated answers into one column, cut out irrelevant columns, then count
# all_time <- gather(answers, key = "Q34_A_ID", value = "Q34_Answer", Q34_Part_1, Q34_Part_2, Q34_Part_3, Q34_Part_4, Q34_Part_5, Q34_Part_6)
# all_time <- all_time[ -c(1:29) ]
# all_time %>% count(Q34_Answer, sort = TRUE)
```

We did not deem this GENERAL SKILL data as useful as originally intended and thus had a stronger case for leaning on Jeff Hale's dataset for this particular area.

We'll expand on the most valuable of software and general skills (later on), but for now **it's worth noting the importance of coding for a data scientist.** As we can see from the above percentages, it plays a large part in a Data Scientists day-to-day life.

#### SURVEY VALUE FILTER

Below is the code for our creation of a table for unique variables related to the VALUE FILTER from the **2018 Kaggle ML & DS Survey** data: 

```{r}
#Education level count
answers %>% count(Q4, sort = TRUE)

##Experience count
answers$Q8 <- as.character(answers$Q8)
answers$Q8[answers$Q8 == "10-May"] <- "5-10"
answers$Q8[answers$Q8 == "15-Oct"] <- "10-15"
answers$Q8[answers$Q8 == "4-Mar"] <- "3-4"
answers$Q8[answers$Q8 == "2-Jan"] <- "1-2"
answers$Q8[answers$Q8 == "3-Feb"] <- "2-3"
answers$Q8[answers$Q8 == "5-Apr"] <- "4-5"

#Storage of corresponding observation counts in "a"
exp_count <- answers %>% count(Q8, sort = TRUE)

#Visualization of years experience count
ggplot(exp_count, aes(x = Q8, y = n)) +
  geom_col(fill = "#00abff") +
  labs(title = "Data Science Survey", subtitle = "Years of Experience") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"), plot.subtitle = element_text(hjust = 0.5)) +
  labs(caption = "2018", x = "", y = "") +
  scale_x_discrete(limits= c("", "0-1", "1-2","2-3", "3-4", "4-5", "5-10", "10-15", "15-20", "20-25", "25-30", "30 +"))

##Income count
mula_count <- answers %>% count(Q9, sort = TRUE)
mula_count
```

From the above tables and chart we can note the different counts for differing educational, experience, and income levels. In particular, we thought the lack of experienced Data Scientists worth emphasizing with our plot. From the *Data Science Survey: Years of Experience* plot we see there are relatively few Data Scientists with 5+ years experience and far fewer with 10+ years experience. 

**It's a young field that's in the midst of defining itself.**

With a clearer understanding of years experience within our dataset, we then set out to explore the education level of our respondents and how their level of education may or may not be related to their income level:

```{r}

answers$Q9 <- as.character(answers$Q9)

income_by_education <- answers %>%
mutate(approx_avg_income = ifelse(Q9 == "0-10,000", "5000",
ifelse(Q9 == "10-20,000", "15000",
ifelse(Q9 == "20-30,000", "25000",
ifelse(Q9 == "30-40,000", "35000",
ifelse(Q9 == "40-50,000", "45000",
ifelse(Q9 == "50-60,000", "55000",
ifelse(Q9 == "60-70,000", "65000",
ifelse(Q9 == "70-80,000", "75000",
ifelse(Q9 == "80-90,000", "85000",
ifelse(Q9 == "90-100,000", "95000",
ifelse(Q9 == "100-125,000", "112500",
ifelse(Q9 == "125-150,000", "137500",
ifelse(Q9 == "150-200,000", "175000",
ifelse(Q9 == "200-250,000","225000",
ifelse(Q9 == "250-300,000", "275000",
ifelse(Q9 == "300-400,000", "350000",
ifelse(Q9 == "400-500,000", "450000",
ifelse(Q9 == "500,000+", "500000", NA)))))))))))))))))))

# unique(income_by_education$Q4)

income_by_education <- income_by_education %>%
mutate(Q4 = ifelse(Q4 == "Bachelor’s degree", "Bachelors",
ifelse(Q4 == "Doctoral degree", "Doctoral",
ifelse(Q4 == "Master’s degree", "Masters",
ifelse(Q4 == "Professional degree", "Professional",
ifelse(Q4 == "No formal education past high school", "High School", ifelse(Q4 == "Some college/university study without earning a bachelor’s degree", "Some college",
ifelse(Q4 == "I prefer not to answer", "No answer", NA))))))))

income_by_education_reporting <- income_by_education %>%
select(Q4, approx_avg_income) %>%
drop_na()

# unique(income_by_education_reporting$Q4)

income_by_education_reporting$approx_avg_income <- as.numeric(income_by_education_reporting$approx_avg_income)

income_by_education_reporting <- subset(income_by_education_reporting, !is.na(Q4))

income_by_education_reporting_graph <- income_by_education_reporting %>%
group_by(Q4) %>%
summarize(approx_avg_income = mean(approx_avg_income))

income_by_education_reporting_graph_w_numbers <- income_by_education_reporting %>%
  mutate(education_years = case_when(Q4 == "Bachelors" ~ 8, Q4 == "Doctoral" ~ 12, Q4 == "High School" ~ 4, Q4 == "Masters" ~ 10, Q4 == "No Answer" ~ 0, Q4 == "Professional" ~ 11, Q4 == "Some college" ~ 6))

income_by_education_reporting_graph_w_numbers <- income_by_education_reporting_graph_w_numbers %>%
  filter(Q4 != "No answer")

# ggplot(income_by_education_reporting_graph, aes(x = Q4, y = approx_avg_income)) +
# geom_col() +
# labs(title = "Data Science Survey", subtitle = "Average Income by Education") +
# theme(plot.title = element_text(hjust = 0.5, face = "bold"), plot.subtitle = element_text(hjust = 0.5)) +
# labs(caption = "2018") +
# theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(income_by_education_reporting_graph_w_numbers, aes(x = education_years, y = approx_avg_income)) +
geom_point(position = "jitter") +
labs(title = "Data Science Survey", subtitle = "Income ~ Education") +
theme(plot.title = element_text(hjust = 0.5, face = "bold"), plot.subtitle = element_text(hjust = 0.5)) +
labs(caption = "2018") +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
geom_smooth(method = "lm") +
  ylim(0,150000) + 
  xlab("Education Years After High School") +
  ylab("Income")

method <- lm(approx_avg_income ~ education_years, income_by_education_reporting_graph_w_numbers)
#summary(method)
method

# p-value of< 2.2e-16 means there is an effect of education years on income
# the r-squared is very small 0.01497 so education doesn't explain much of 
# the overall variability in income
```
What we found, and what can be seen above is that there is a positive relationship between education and income, as painted by the *blue line* fit to the plotted data. Thus, the higher a Data Scientist's level of education, the higher their potential earnings.

With a clearer picture of experience, education, and income levels, we then set out to build our "value filter". We elected to filter observations / respondents with (1) income >= $100,000 and (2) 10+ yrs experience. 

Our aim was to extract the "perspective" of those that are most experienced and highest income earning (regardless of educational background) to see whether their perspective was consistent with or different from that of the general survey populace:

```{r}
##Filter data for (1) income >= $100,000, and 10+ yrs experience
answers <- answers %>% 
  mutate(value_filter = ifelse((answers$Q8 == "10-15" | answers$Q8 == "15-20" | answers$Q8 == "20-25" | answers$Q8 == "25-30" | answers$Q8 == "30 +") & (answers$Q9 == "100-125,000" | answers$Q9 == "125-150,000" | answers$Q9 == "150-200,000" | answers$Q9 == "200-250,000" | answers$Q9 == "250-300,000" | answers$Q9 == "300-400,000" | answers$Q9 == "400-500,000" | answers$Q9 == "500,000+"), "yes", "no"))

#Thus, extracting 728 observations from more than 23,000
filtered <-answers[ which(answers$value_filter =="yes"),]

```

We then applied our value filter to see whether there were any observable differences between the general survey populace and those that have 10+ years experience and are earning $100k+ more per year. In short, *do those with more experience and market-value see things differently or similarly?*

Below are the tables related to SOFTWARE SKILLS from the **2018 Kaggle ML & DS Survey** data **with our value filter applied**: 

```{r}
#Filtered "regular basis" software (Q16) count
##Gather associated answers into one column, cut out irrelevant columns, then count
#Note: observation count = 16 cols x 748 observations (in filtered)
f_reg <- gather(filtered, key = "Q16_A_ID", value = "Q16_Answer", Q16_Part_1, Q16_Part_2, Q16_Part_3, Q16_Part_4, Q16_Part_5, Q16_Part_6, Q16_Part_7, Q16_Part_8, Q16_Part_9, Q16_Part_10, Q16_Part_11, Q16_Part_12, Q16_Part_13, Q16_Part_14, Q16_Part_15, Q16_Part_16)
f_reg <- f_reg[ -c(1:20) ]
f_reg <- f_reg %>% count(Q16_Answer, sort = TRUE)%>% rename(f_reg = n)
f_reg <- f_reg [-1,] #remove empty row
f_reg

#Filtered "use most often" software (Q17) count
f_mo <- filtered %>% count(Q17, sort = TRUE)%>% rename(f_mo = n)
f_mo <- f_mo[-2,] #remove empty row
f_mo

#Software language "recommend to aspiring DSs" (Q18) count
f_rec <- filtered %>% count(Q18, sort = TRUE)%>% rename(f_rec = n)
f_rec <- f_rec[-3,] #remove empty row
f_rec

```

We can extend a number of observations from these "value filtered" tables to those of the earlier general data set:

* Python has the highest rating for regular / most often (Q16, Q17) use as well as the highest level of recommendation to aspiring data scientists (Q18). 
* SQL is #2 for regular use (Q16) and R is #2 for most often and highest level of recommendation to aspiring data scientists (Q18)
* The top 3 languages were consistently Python, R, and SQL.

The disproportionate level of Python use and recommendation is explored in more detail later. Prior to moving on to software and general skills though, we applied a second value filter to our general survey data to explore the effect of education on income:

--------------------------------------------------------------------------------

#### JOB BOARD SKILLS

We put Jeff Hale’s in a form to more easily compare to / pull from by gathering counts for unique variables related to SOFTWARE SKILLS and then GENERAL SKILLS:
```{r}
#Consolidate job board data (ie. LinkedIn v Indeed) for SOFTWARE SKILLS
tech_skills[ , c(2,3,4,5)]  <- apply(tech_skills[ , c(2,3,4,5)], 2, function(x) as.numeric(str_remove(x, ",")))
tech_skills_final <- tech_skills%>%
  filter(Keyword != "Total", LinkedIn.. != "")%>%
  select(Keyword:Monster)%>%
  mutate(total = rowSums(. [2:5]))%>%
  arrange(desc(total))
##Then sum all rows to get all job search results per keyword and re-plot

#Consolidate job board data (ie. LinkedIn v Indeed) for GENERAL SKILLS
genl_skills[ , c(2,3,4,5)]  <- apply(genl_skills[ , c(2,3,4,5)], 2, function(x) as.numeric(str_remove(x, ",")))
genl_skills_final <- genl_skills%>%
  select(Keyword:Monster)%>%
  filter(Keyword != "Total", LinkedIn != "")%>%
  mutate(total = rowSums(. [2:5]))%>%
  arrange(desc(total))

#Subset data for most popular observations: visualization, statistics, mathematics, machine learning, deep learning, computer science, communication, analysis
genl_skills_final <- genl_skills_final %>% filter(total > 3000)
```

--------------------------------------------------------------------------------

### VISUALIZE & ANALYZE

While the tables and plots upto this point helped tell the story of our exploration of the **2018 Kaggle ML & DS Survey** and **The Most in Demand Skills for Data Scientists** datasets, the plots that follow directly answer the question at hand.

First, we set out to gain insight as to what the most valuable GENERAL SKILLS might be. We pulled on Jeff Hale's data set and plot for total counts (across job boards) for the most common key words. We then plotted for the Top 10 General Skills:

```{r}
#Plot for general skills from Jeff Hale's dataset

ggplot(genl_skills_final) +
  geom_bar(aes(x = Keyword , y = total, fill = Keyword), stat = "identity", position = "dodge", width = 1) + coord_flip() +
  theme(legend.position = "none") +
  labs( title = "Job Board General Skills Mentions", x = "", y = "", fill = "Source")

```

The top 10 skills include: *visualization, statistics, NLP composite, mathematics, machine learning, deep learning, computer science, communication, analysis, and AI composite.* This subset of skills includes some soft skills (ie. communication) but mostly technical skills (ie. statistics). 

When we look at the top 3 "general" skills (analysis, machine learning, and statistics), they're all arguably technical. Thus, it would seem the most valuable skills for a Data Scientist would be a very strong technical toolbox with the ability to glean the "golden nuggets" and telling visualizations from the data at hand and relay them to the rest of their team / management / the customer / etc.

From the plot above, we get the idea that the role is highly technical and from the earlier plot of % daily activities coding, we saw that a large portion of a Data Scientist's day is spent coding. Thus, it's very valuable to know how to program. And with that in mind, there are certain programming languages that are more valuable and applicable in the realm of Data Science than others.

So with our next visualization, we set out to gain insight into what the most valuable SOFTWARE SKILLS might be. We pulled on Jeff Hale's data set and plot for total counts (across job boards) for the most common key words. We then plotted for the Top 8 Software Skills:

```{r}
keyword_comp <- tech_skills_final %>%
  inner_join(rec,c("Keyword" = "Q18"))%>%
  left_join(reg,c("Keyword" = "Q16_Answer"))%>%
  left_join(mo,c("Keyword" = "Q17"))%>%
  pivot_longer(cols = LinkedIn:mo, names_to = "source" , values_to = "frequency" )%>% 
      group_by(source) %>%
      mutate(pct  = frequency / sum(frequency))

keyword_comp%>%
  filter(source == "total" | source == "rec")%>%
  ggplot() +
  geom_bar(aes(x = Keyword , y = pct, fill = source), stat = "identity", position = "dodge", width = 1)+
  scale_fill_manual(labels = c("Data Scientist", "Job Board"), values = c("#9999CC", "#66CC99")) +
  labs( title = "Data Scientist Recommendations vs. Job Board", x = "Software Language", y = "Proportion of Mentions", fill = "Source")+
  coord_flip()

```

The top 8 languages include: *SQL, Scala, SAS, R, Python, Javascript, Java, and C++*. And where the bottom 5 languages had varying proportion of mentions (depending on whether it was drawn from the Job Board data or from the Survey data), **the top 3 software skills were consistently: Python > R > SQL.**

What's interesting to note is the discrepancy between what Data Scientists perceive as most important (the languages they most recommend) v. what the frequency of corresponding software skills as keywords in job postings. 

To expand on this point ... if we were to look at the survey data, we'd see that nearly 80% of respondents recommend Python, whereas this value is just below 30% when we look at corresponding job board proportion of mentions. Thus, the perceived value of Python may be much greater than its actual value, whereas the perceived value of languages like R and SQL may be lower than their actual value.

Prior to concluding our analysis, we deemed it worthwhile to re-visualize the aforementioned plot *without Python* to more clearly see representative proportions of the rest of the languages:

```{r}
keyword_comp%>%
  filter(source == "total" | source == "rec", Keyword != "Python")%>%
  ggplot() +
  geom_bar(aes(x = Keyword , y = pct, fill = source), stat = "identity", position = "dodge", width = 1)+
  scale_fill_manual(labels = c("Data Scientist", "Job Board"), values = c("#9999CC", "#66CC99")) +
  labs( title = "Data Scientist Recommendations vs Job Board", subtitle = "(Without Python)", x = "Software Language", y = "Proportion of Mentions", fill = "Source") +
  coord_flip()
```
What we see from the plot above is that the 7 "runner-up" most popular languages all carry a disproportionate Proportion of Mentions when compared to the associated survey bar.

We see that R and SQL run 1-2 for their popularity in share of mentions (which we knew before), but then the picture of the next tier of languages clarifies quite a bit: *Java, SAS, Scala, C++, and Javascript* are all worth "honorable mention". Javascript may be more niche with less than 2.5% Proportion of Mentions on job postings but Java is nearly 10% which is well worth considering if you're an aspiring Data Scientist or a seasoned Data Scientist looking to sharpen your arsenal. 

--------------------------------------------------------------------------------

### CONCLUDE

We return to the question at hand:

> What are the most valued data science skills?

In "twice sifting" our data and adapting our approach based on insights gained, we pulled from both the **2018 Kaggle ML & DS Survey** and **The Most in Demand Skills for Data Scientists** data, to determine that the most valued data science skills cover two realms: general skills and software skills.

The analysis portion goes more in-depth regarding these realms. For sake of clarity and conciseness, we’ll take an Olympic medal approach (think: gold, silver, bronze) and note *only* the top three in each respective category below.

* **General skills**: analysis > machine learning > statistics.

* **Software skills**: Python > R > SQL.

Although more comprehensive or recent of data could be used to verify, our multiple sifts and plots of the data confirm not only the skills listed above but the fact that there is a discrepancy between perceived and real value. 

When we take into account the proportion of mentions for Job Board key words v. survey responses, it seems that Data Scientists over-value the importance of Python while under-valuing the importance of R and SQL. Thus, as a Data Scientist, there may be opportunity in placing more value on these languages.

--------------------------------------------------------------------------------

### NOTE FROM THE TEAM

Our team attempted to collaborate remotely through duration of this project, and although it was a valiant attempt, one take-away we all agreed upon was the importance of using better real-time collaborative tools (ie. repl.it or git) rather than copy-pasting code between Githubs and/or Slack-and-Github. 

After leaning too heavily on manual entry (and thus unnecessarily losing time and amplifing the possibility of error), our motivation for incorporating these tools and methods into future projects is clear as day :)