---
title: "DATA 607 Project 2 - Dataset 1"
author: "Magnus Skonberg"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        number_sections: false
        theme: cerulean
        highlight: tango
        font-family: "Arial"
---

```{r load-packages, include=FALSE}
library(tidyverse)
library(openintro)
library(dplyr)
library(tidyr)
library(RCurl)
library(ggplot2)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Background

The purpose of this assignment is to *tidy and transform data*. The dataset of interest describes **Happiness v GDP for the United States and Finland** and our task is to: 

(1) Create a .csv file in a "wide" structure.
(2) Read this .csv and use tidyr and dplyr to tidy and transform.
(3) Analyze Happiness and GDP between the nations.
(4) Provide the .rmd file, rpubs link, and descriptions of steps taken.

--------------------------------------------------------------------------------

\clearpage

### (1) Create a .csv.

The original data set was downloaded from Kaggle and is [available here.](https://www.kaggle.com/unsdsn/world-happiness)

The Happiness Report uses data from the Gallup World Poll and is based on answers to the main life evaluation questions asked in the poll. Rankings run from 0 (worst) to 10 (best), with each individual responding based on the state of their current life v. their ideal life within the nation they call home.

A wide format .csv was created per assignment description and [uploaded to github.](https://github.com/Magnus-PS/CUNY-SPS-DATA-607/blob/Project-2/happy_data.csv)

Our dataset was pulled from the larger Kaggle dataset and focuses exclusively on the United States (our home nation) v Finland (the happiest nation in 2019). *Our analysis will compare happiness and GDP data between these nations.*

With the .csv available on github, we shift to reading from this file *prior to* tidying and transforming the data therein.

### (2) Read, tidy, and transform.

We read the .csv (in its raw form) from github and store corresponding data in a variable named "data" (I know ... very creative).

```{r}
#Get URL, read .csv (in raw form) from github, and put into tabular form
url <- getURL("https://raw.githubusercontent.com/Magnus-PS/CUNY-SPS-DATA-607/Project-2/happy_data.csv")
data <- read.csv(text = url)
data <- as_tibble(data)

#Show what we're working with:
data

```
Once we've read from .csv and stored our data, we shift to tidying and transforming the data therein.

To start, we tidy:

```{r}
#Tidy / shape our data (using tidyr)

##Provide fitting column names for the 1st few metrics so that we can read them on the graph later.
names(data)[names(data) == "Happiness.Score"] <- "Happiness"
names(data)[names(data) == "GDP.per.Capita"] <- "Wealth"
names(data)[names(data) == "Social.Support"] <- "Fellowship"
names(data)[names(data) == "Healthy.life.expectancy"] <- "Health"

##Make observations from variables using gather()
##This is also when we convert from a 'wide' to 'long' format
data_long <- gather(data, "Metric", "Score", 3:7, factor_key=TRUE)

##Sort 'Country' column alphabetically
data_long <- data_long[order(data_long$Country),]

##Make variables from observations using spread()
data_long <- spread(data_long, "Year", "Score")
data_long

#We plot all metrics for 2015:
qplot(x=Metric, y=`2015`, data=data_long, col=Country, main="2015 Happiness v GDP Metrics", xlab="Metric", ylab="Value")

#We plot all metrics for 2019:
qplot(x=Metric, y=`2019`, data=data_long, col=Country, main="2019 Happiness v GDP Metrics", xlab="Metric", ylab="Value")

```

What started as a *less clear* 4x7 'wide' table ("data") is now a  *tidy* 10x4 'long' table ("data_long") with plots for 2015 and 2019 showing the changes in metrics for each nation.

What we can see, when we flip between plots is that Finland edges the US in 4/5 metrics. More specifically, from 2015 to 2019, Finland put tremendous ground between itself and the US for Happiness, the US kept the lead in terms of Wealth, Freedom metrics remained close, and for Fellowship and Health Finland gained a little bit of ground on the US.

...

With tidying complete, we move on to transforming our data. We'll make use of dplyr() to transform data and perform mathematical calculations to solve for our variables. After this point, we should have all we need to *clearly* observe the difference in Happiness v GDP for Finland and the US. More specifically, **how each nation's Happiness v GDP changed between 2015 and 2019.**


At a high-level, we have *two* variables of interest: GDP and Happiness. The other *three* variables may play some role in the metric but that is not our focus and thus we will drop these Metrics from our consideration, and take the change in Happiness over the change in GDP for each Finland and the US.

*(Happiness (2019) - Happiness (2015)) / (GDP (2019) - GDP (2015))*

We can then compare and plot these values to see what affect, if any income has on happiness in 2019's "Happiest Country in the World" (Finland) v. our home nation (the US).

```{r}
#Transform the data: dplyr

##Add a column showing the change from 2015 to 2019
data_long$Chg <- data_long$`2019` - data_long$`2015`

##Extract data for Happiness and Wealth
hw <- filter(data_long, Metric == "Happiness" | Metric == "Wealth")

hw <- select(hw, Country, Metric, Chg)
hw

##Plot the change in Happiness / change in Wealth for each nation to see if we might draw any interesting insight.
fin_h_chg <- as.numeric(hw[1,3])
fin_w_chg <- as.numeric(hw[2,3])
fin_val <- round(fin_h_chg / fin_w_chg, 3)

us_h_chg <- as.numeric(hw[3,3])
us_w_chg <- as.numeric(hw[4,3])
us_val <- round(us_h_chg / us_w_chg, 3)


##Plot the average score v parent education level for Math
barplot(c(fin_val, us_val), beside=TRUE, main = "Happiness v Wealth for the Finland and the US", xlab = "Country", ylab = "Happiness/Wealth", ylim = c(-10, 10), names.arg = c("Finland", "United States"), col=c("#6699FF", "#6699FF"), border="white")

text(0.75, 6.5, as.character(fin_val))
text(1.90, -5, as.character(us_val))

```

###LEFT OFF HERE - explain methodology, analyze results then conclude :)