---
title: "DATA 607 - Wk3 HOL"
author: "Magnus Skonberg"
date: "9/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
```

### 1. Create a vector of strings and identify strings that match a particular criteria.

```{r}
fish <- c("salmon", "tilapia", "cod", "tuna", "swordfish" )
fish == "salmon"
```

### 2. Replace instances of a string with another of your choosing.

```{r}

grep(pattern = 'til', fish, value = TRUE, ignore.case = TRUE)
gsub(pattern = "til", replacement = "shark", fish[2])

```

### 3. Identify positions of interest within your string.

```{r}
which(fish == "cod")
```

### 4. Use separators to combine multiple string outputs together.
```{r}
str_c(fish[1], fish[2], sep = ", ")
```

### 5. Use collapse to combine multiple string outputs together.

```{r}

str_c(c(fish[1], fish[2], fish[3]), collapse = ", ")

```

### 6. Change the case of strings.

```{r}
str_to_upper(fish[1])
str_to_lower(fish[1])
```
### 7. Convert dates between different formats.

```{r}
now()

sept_ymd <- ymd("2020-09-08")

mdy("September 8th, 2020")

dmy("8-Sep-2020")

ymd(20200908)

as_datetime(today())

as_date(now())

year(now())

datetime <- ymd_hms("2016-07-08 12:34:56")

year(datetime) <- 2020
hour(datetime) <- hour(datetime) + 1

datetime


```
### 8. Identify how long ago dates occurred.

```{r}

m_age <- today() - ymd(19910528)
m_age / 365

next_year <- today() + years(1)
(today() %--% next_year) %/% days(1)


```